Fast solvers for partial differential equations \\ subject to inequalities

### 2023-11-09 talk-dms/slides.tex
inner cone approx figure
what I practice on

\subsection{earlier approaches}

@@

\begin{itemize}
\item on simpler VI problems, \alert{nested iteration} helps
    \begin{itemize}
    \item[$\circ$] nested-iteration = grid-sequencing
    \item[$\circ$] on classical obstacle problem: grid-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
    \end{itemize}
\item not as effective on harder (\emph{lower-regularity, complicated free boundary}) problems
\end{itemize}

@@

\begin{itemize}
\item to explain ``constraint decomposition'' in FAS\underline{CD}, start with subspace decomposition in the multilevel case
\item assume mesh levels are nested: $\Omega^j \subset \Omega^{j+1}$ nodewise
\item let $\mathcal{V}^j$ be the space of FE functions on $\Omega^j$
    \begin{itemize}
    \item[$\circ$] standard FE spaces: \, $\mathcal{V}^j \subset \mathcal{V}^{j+1}$
    \end{itemize}
\begin{block}{definition}
$\ds \mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ \quad is called a \emph{subspace decomposition} (Xu 1992)
\end{block}
    \begin{itemize}
    \item[$\circ$] \emph{non}-unique vector space sum
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs extends subspace decomposition to convex subsets
    \begin{itemize}
    \item[$\circ$] non-trivial
    \end{itemize}
\item suppose $\mathcal{K} \subset \mathcal{V}$ is a closed and convex admissible subset
\item for a subspace decomposition $\mathcal{V} = \sum_i \mathcal{V}_i$, write the admissible subset as a sum
    $$\mathcal{K} = \sum_i \mathcal{K}_i$$
where $\mathcal{K}^i \subset \mathcal{V}^i$, with projections $\Pi_i : \mathcal{K} \to \mathcal{K}_i$
\end{itemize}
\end{frame}

@@

@@

\begin{itemize}
\item recall $\mathcal{K} = \{v \ge \psi\}$ in classical obstacle problem
\item define \alert{defect obstacle} for a fine-level iterate $w^J$:
    $$\chi^J = \psi^J - w^J$$
\item \alert{monotone restriction} generates obstacles on each level:
	$$\chi^j = R^{\oplus} \chi^{j+1}$$
\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \chi^j - \chi^{j-1}\}$
\item get CD of fine-level constraint set:
	$$\mathcal{U}^J = \sum_{i=0}^J \mathcal{D}^i \phantom{KLDFJSKJDS SDF}$$
\item multiplicative CD iteration $\to$ V-cycle
\end{itemize}

\vspace{-25mm}
\hfill \includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
\end{frame}

@@

% case of lower obstacle only
\begin{pseudo}
FIXME
\end{pseudo}
\end{frame}

>>>>

\subsection{an earlier approach}

@@

\begin{itemize}
\item on simpler VI problems, \alert{nested iteration} helps
    \begin{itemize}
    \item[$\circ$] nested-iteration = grid-sequencing
    \item[$\circ$] on classical obstacle problem: mesh-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
    \end{itemize}
\item not as effective on harder (\emph{lower-regularity, complicated free boundary}) problems
\end{itemize}

@@

\begin{itemize}
\item to explain the ``CD'' (constraint decomposition) in FAS\underline{CD}, start with subspace decomposition in the multilevel case
\item assume mesh levels are nested: $\Omega^j \subset \Omega^{j+1}$ nodewise
\item let $\mathcal{V}^j$ be the space of FE functions on $\Omega^j$
    \begin{itemize}
    \item[$\circ$] standard FE spaces: \, $\mathcal{V}^j \subset \mathcal{V}^{j+1}$
    \end{itemize}
\begin{block}{definition}
$\ds \mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ \quad is called a \emph{subspace decomposition} (Xu 1992)
\end{block}
    \begin{itemize}
    \item[$\circ$] \emph{non}-unique vector space sum
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Tai's (2003) constraint decomposition for VIs extends subspace decomposition to convex subsets
    \begin{itemize}
    \item[$\circ$] extends non-trivially
    \end{itemize}
\item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
\item assume $\mathcal{V}^J = \sum_i\mathcal{V}^i$ is a subspace decomposition
\begin{block}{definition}
$\ds \mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$ \quad is a \emph{constraint decomposition} (CD) if there are closed and convex subsets $\mathcal{K}^i\subset \mathcal{V}^i$, and (nonlinear) projections $\Pi_i : \mathcal{K}^J \to \mathcal{K}^i$, so that $\ds v = \sum_{i=0}^J \Pi_i v$ and a stability condition applies (not shown)
\end{block}
\end{itemize}
\end{frame}

\begin{itemize}
\item observation: generally $\mathcal{K}^i \not\subset \mathcal{K}^J$
\item picture below: obstacle problem on a two-point mesh ($\mathcal{V} \cong \RR^2$)
\end{itemize}

\bigskip
\begin{center}
\includegraphics[width=0.6\textwidth]{../paper/genfigs/cartoon.pdf}
\end{center}
\end{frame}

@@

\begin{frame}{defect obstacles}

\begin{itemize}
\item recall $\mathcal{K} = \{v \ge \psi\}$ in classical obstacle problem
\begin{block}{definition}
the \emph{defect obstacle} for a fine-level iterate $w^J \in \mathcal{K}^J = \{v\,:\,v\ge \psi^J\}$ is an FE function in $\mathcal{V}^J$:
    $$\chi^J = \psi^J - w^J$$
\end{block}
\item we generate the CD through defect obstacles $\chi^j$ on each level
\item via \emph{monotone restriction}: \quad $\ds \chi^j = R^{\oplus} \chi^{j+1}$
\end{itemize}

\begin{center}
\input{tikz/Rplusminus.tex}
\end{center}
\end{frame}

\begin{frame}{down and up sets}

\begin{itemize}
\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \phi^j=\chi^j - \chi^{j-1}\}$
\item get CD of fine-level constraint set:
	$$\mathcal{U}^J = \sum_{i=0}^J \mathcal{D}^i \phantom{KLDFJSKJDS SDF}$$
\end{itemize}

\begin{center}
\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
\end{center}
\end{frame}

@@

\centering
\includegraphics[width=0.7\textwidth]{figs/innerconeapprox.png}
\end{frame}

@@

% case of lower obstacle only
\begin{pseudo}
\pr{fascd-vcycle}(J,\ell^J,\psi^J;w^J)\text{:} \\+
    $\chi^J = \psi^J - w^J$ \\
    for $j=J$ downto $j=1$ \\+
      $\chi^{j-1} = \maxR \chi^j$ \\
      $\phi^j = \chi^j - P\chi^{j-1}$ \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$  \hspace{1.0cm} \ct{in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$ \hspace{1.0cm} \ct{in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \hspace{1.0cm} \ct{in $\mathcal{U}^j$} \\-
    return $w^J+z^J$
\end{pseudo}
\end{frame}
